---
title: "PlantBug_Markdown"
author: "MJL"
date: "10/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Plant bug network analyses Fall 2018
Our goals for this project are to:
 1. explore the network across the different Hemiptera species in the Museum database  
 2. better understand the geographical and phylogenetic patterns of these networks

### Load in necessary libraries for data analysis of Plant-Bug networks:  
 - ggplot2  
 - RColorBrewer  
 - ggthemes  
 - bipartite  
 - tidyverse
 - dplyr
```{r message = FALSE, warning = FALSE}
library(ggplot2)
library(RColorBrewer)
library(ggthemes)
library(bipartite)
library(tidyverse)
library(dplyr)
library(plyr)
```


### Load in necessary datasets
As of November 2018, we'll be using the "hostNetwork_copy.tsv" that Katja pulled and put together from the museum data (included in this package).
```{r}
# Read in data file and read as data frame.
data <- read.delim("../data/hostNetwork.tsv", header = TRUE)

# Create new columns for full species names of both the host species and the insect species.
data$host <- paste(data$h_genus, data$h_species)
data$insect <- paste(data$i_genus, data$i_species)



#new table with columns
#write.table(data, "collectionsConservationBio/temp-data/data.txt", na = "NA", row.names = FALSE, col = TRUE, append = FALSE, sep="\t", quote=FALSE)

#creating a sublist of data where host plants and host insects are redlisted
#subdata <- subset(data, redList == TRUE)
```




### 1a. Creating webs and making calculations

#### _Webs and calculations with bipartite_
Here we're filtering the big dataset to host family level and then creating the interaction network/calculating network metrics using the package, bipartite. 

We focus on connectance, weighted connectance, interaction evenness, and nestedness.
```{r}
### Functions

## dataframe to matrix
web <- function(df) {
  as.data.frame(frame2webs(df, varnames = c("host", "insect", "h_family", "coll_number_same_h"), type.out = 'array'))
}

## multiple network metric calculations
multi_metric <- function(df) {
  c(
    # connectance
    connect <- networklevel(df, index = 'connectance'),
    # weighted connectance
    weight_connect <- networklevel(df, index = 'weighted connectance'),
    # interaction evenness
    even <- networklevel(df, index = 'interaction evenness'),
    # nestedness
    nest <- networklevel(den_web, index = 'nestedness')
)}


```



Size bins: 0-25
```{r}
# Dennstaedtiaceae <- 16 obs
den <- data %>%
  filter(h_family == "Dennstaedtiaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
# transform observation data frame into an network matrix array
multi_metric(web(den))

### Ginkgoaceae <- only 3 observations
gin <- data %>%
  filter(h_family == "Ginkgoaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(gin))

### Aspleniaceae <- 1 obs
asp <- data %>%
  filter(h_family == "Aspleniaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(asp))
```

Size bin: 25-50
```{r}
# Lythraceae <- 41 obs
lyt <- data %>%
  filter(h_family == "Lythraceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(lyt))

# Primulaceae <- 33 obs
pri <- data %>%
  filter(h_family == "Primulaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(pri))

# Cycadaceae <- 31 obs
cyc <- data %>%
  filter(h_family == "Cycadaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(cyc))
```

Size bin: 50-75
```{r}
# Elaeagnaceae <- 54 obs
ela <- data %>%
  filter(h_family == "Elaeagnaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(ela))

# Cucurbitaceae <- 55 obs
cuc <- data %>%
  filter(h_family == "Cucurbitaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(cuc))

# Liliaceae <- 55 obs
lil <- data %>%
  filter(h_family == "Liliaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(lil))
```

Size bin: 75-100
```{r}
# Magnoliaceae <- 82 obs
mag <- data %>%
  filter(h_family == "Magnoliaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(mag))

# Casuarinaceae <- 76 obs
cas <- data %>%
  filter(h_family == "Casuarinaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(cas))

# Chenopodiaceae <- 93 obs
che <- data %>%
  filter(h_family == "Chenopodiaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(che))
```



```{r warning = FALSE}
### Next, we'll use the Euphorbiaceae family which has 207 observations. Slightly larger, but still manageable.
eup <- data %>%
  filter(h_family == "Euphorbiaceae") %>% 
  select(host, insect, h_family, coll_number_same_h)
multi_metric(web(eup))
```

#### Make a more streamlined function to go through families 
```{r}
# 1. take groups of families (only those with n < 200?)
family_groups <- data %>%
  group_by(h_family) %>% 
  filter(length(h_family) < 100) %>% 
  arrange(h_family) %>% 
  select(h_family, host, i_family, insect, coll_number_same_h)

# 2. create interaction matrix
# 3. convert matrix to data frame
connect <- function(df) {
networklevel(as.data.frame(frame2webs(df, varnames = c("host", "insect", "h_family", "coll_number_same_h"), type.out = 'array')), index = "connectance")
}

# 4. calculate matrix metrics
connect_fam <- dlply(family_groups, family_groups$h_family, .fun = connect)

# 5. print in metric data frame
  # a. connectance
  # b. weighted connectance
  # c. interaction evenness
  # d. nestedness


```



#### _Analyses of calculations_
Following the above network calculations, we are now going analyze general trends in network metrics. 

```{r}
# Filter data for histogram looking at the number of unique interactions per host family
fam_hist_data <- data %>%
  group_by(h_family) %>% 
  filter(length(h_family) >100) %>% 
  select(h_family, host, i_family, insect, coll_number_same_h)
hostfam_hist <- ggplot(fam_hist_data, aes(x = coll_number_same_h)) +
  geom_histogram()
hostfam_hist


# check normality of metric spreads
net_mets <- read_csv("plantbug_selectNetMets.csv") #%>% 
  #filter(h_family != "Euphorbiaceae")

hist_connect <- ggplot(net_mets, aes(x = connectance)) +
  geom_histogram()
hist_connect


## basic correlation between each metric and family size

# connectance and family size
connect_famsize <- ggplot(net_mets, aes(x = h_fam_size, y = connectance)) +
  geom_point()
connect_famsize

# weighted connectance and family size
wconnect_famsize <- ggplot(net_mets, aes(x = h_fam_size, y = net_mets$weigh_connect)) +
  geom_point()
wconnect_famsize

# relative specialization of insect families
```

### Interaction network for just host and insect families
```{r}

```



### 1b. Do these analyses change if we look from the insect family perspective?