---
title: "Plant Bug Network Analyses"
author: "K. Seltmann & M. Lee"
date: "14-Dec-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Plant bug network analyses Fall 2018
Our goals for this project are to:
 1. explore the network across the different Hemiptera species in the Museum database  
 2. better understand the geographical and phylogenetic patterns of these networks

### Load in necessary libraries for data analysis of Plant-Bug networks:  
```{r message = FALSE, warning = FALSE}
library(bipartite)
library(tidyverse)
library(dplyr)
library(OneR)
library(mltools)
```


### Load in necessary datasets
As of December 2018, we'll be using the "lat-long-Miridae.txt" that Katja pulled and put together from the museum data (included in this package).
```{r}
# Read in data file and read as data frame. Beware, this dataset is LARGE.
  # 282,718 observations and 24 columns

data1 <- read.delim("~/Desktop/PlantBugs/collectionsConservationBio/PlantBugs-2017/temp-data/lat-long-Miridae.txt", header = TRUE)

# filter out bad lat/long
data <- data1 %>% 
  filter(dlong < "0")


```

# 1. BASIC DATA VISUALIZATION

### _network nodes_
```{r}
######### MIRIDAE IS 1 FAMILY #########
### 362 unique Miridae genera
num_gen_mirid <- length(unique(data$i_genus))
num_gen_mirid
### 2544 unique Miridae species
num_spp_mirid <- nrow(unique(data[, c("i_genus","i_species")]))
num_spp_mirid
######################################


######### PLANT INFORMATION #########
### 156 unique plant families
num_fam_host <- length(unique(data$h_family))
num_fam_host

### DISCREPANCY--unique plant genera: either 745 or 806
# unique combinations of family and genus gives 806
num_gen_host <- nrow(unique(data[, c("h_family","h_genus")]))
num_gen_host
# unique genera gives 745
num_gen_host2 <- length(unique(data$h_genus))
num_gen_host2

### DISCREPANCY--unique plant species
# unique host family, genus, species combinations
num_spp_host <- nrow(unique(data[, c("h_family","h_genus","h_species")]))
num_spp_host
# unique species IDs
num_spp_host2 <- length(unique(data$h_species_id))
num_spp_host2
#####################################
```


### _spread of genus data_
```{r}
# all mirid genera and the number of times they are in the dataset
# thus this represents the number of observations there are for each genus
gen_bar <- ggplot(data, aes(x = i_genus)) +
  geom_bar() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
        ) +
  xlab("Mirid Genera") +
  ylab("Observation Events")
gen_bar

# log of counts
log_gen_bar <- ggplot(data, aes(x = i_genus)) +
  geom_bar() +
  scale_y_log10() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
        ) +
  xlab("Mirid Genera") +
  ylab("Observation Events (log transformed)")
log_gen_bar


# subset of mirid genera with A LOT of observations
large_genera <- data %>% 
  group_by(i_genus) %>% 
  filter(length(i_genus) > 2500)

# new bar graph with only the genera with the most observations
# then colored by species
biggen_bar <- ggplot(large_genera, aes(x = i_genus)) +
  geom_bar(aes(fill = i_species), show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("Mirid Genera with over 2500 Observations") +
  ylab("Observation Events")
biggen_bar


```

### _latitude/longitude visualizations_
```{r}
### lat/long visualization
lat_long_scatter <- ggplot(data, aes(x = dlong, y = dlat)) +
  geom_point(aes(colour = i_genus), show.legend = FALSE) +
  xlab("Longitude (decimal degrees)") +
  ylab("Latitude (decimal degrees)")
lat_long_scatter

### lat/long visualization for just CA
CA_latlong <- data %>%
  filter(stateProv == "California" & dlat != "0")
cal_scatter <- ggplot(CA_latlong, aes(x = dlong, y = dlat)) +
  geom_point(aes(colour = i_genus), show.legend = FALSE) +
  xlab("Longitude (decimal degrees)") +
  ylab("Latitude (decimal degrees)")
cal_scatter

```


# 2. PRELIMINARY ANALYSES OF BINNED GENERA

### scatter plot: # plants interacted with by # species in genus
```{r}
# summary of how many insect species and host families per host genus
summary_gen <- data %>%
  group_by(i_genus) %>% 
  summarise(spp_pergen = length(unique(i_species_id)), host_pergen = length(unique(h_family)))

# scatter relationship of number of species per insect genus by the number of plant families
pergenus_scatter <- ggplot(summary_gen, aes(x = spp_pergen, y = host_pergen)) +
  geom_point() +
  xlab("Insect species per genus") +
  ylab("Host families per genus") +
  stat_smooth(method = "lm", formula = y ~ log(x))
pergenus_scatter
```

### binning continuous data
```{r}
# binning continuous data 
dat_binned <- data %>% 
  mutate(
    elev_bin = bin_data(ElevM, bins = 30),
    long_bin = bin_data(dlong, bins = 30),
    lat_bin = bin_data(dlat, bins = 30)
  )
```


### relationship between longitude and # plant families interacted with
```{r}
# group_by above data and then count plant family interactions
long <- dat_binned %>% 
  group_by(long_bin) %>% 
  summarise(host_perlong = length(unique(h_family)))
# bar graph of long data and plant family interactions
long_scatter <- ggplot(long, aes(x = long_bin, y = host_perlong)) +
  geom_point() +
  xlab("Longitude (decimal degrees)") +
  ylab("Host families") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
long_scatter

# same of figure above without labels 
long_no_labels <- ggplot(long, aes(x = long_bin, y = host_perlong)) +
  geom_point() +
  xlab("Longitude (decimal degrees)") +
  ylab("Host families") +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
long_no_labels
```


### relationship between latitude and # plant families interacted with
```{r}
# group_by above data and then count plant family interactions
lat <- dat_binned %>% 
  group_by(lat_bin) %>% 
  summarise(host_perlat = length(unique(h_family)))
# bar graph of lat data and plant family interactions
lat_scatter <- ggplot(lat, aes(x = lat_bin, y = host_perlat)) +
  geom_point() +
  xlab("Latitude (decimal degrees)") +
  ylab("Host families") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
lat_scatter

# same of figure above without labels 
lat_no_labels <- ggplot(lat, aes(x = lat_bin, y = host_perlat)) +
  geom_point() +
  xlab("Latitude (decimal degrees)") +
  ylab("Host families") +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
lat_no_labels
```


### relationship between elevation and # plant families interacted with
```{r}
# group_by above data and then count plant family interactions
elev <- dat_binned %>% 
  group_by(elev_bin) %>% 
  summarise(host_perelev = length(unique(h_family)))
# bar graph of lat data and plant family interactions
elev_scatter <- ggplot(elev, aes(x = elev_bin, y = host_perelev)) +
  geom_point() +
  xlab("Elevation (m)") +
  ylab("Host families") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
elev_scatter

# same of figure above without labels 
elev_no_labels <- ggplot(elev, aes(x = elev_bin, y = host_perelev)) +
  geom_point() +
  xlab("Elevation (m)") +
  ylab("Host families") +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
elev_no_labels
```







# 3. NETWORK CREATION
### Interaction network for just host and insect families
```{r}
family_groups <- data %>%
  group_by(h_family) %>% 
  filter(length(h_family) < 100) %>% 
  arrange(h_family) %>% 
  select(h_family, h_genus, host, i_family, i_genus, insect, coll_number_same_h)

family_web <- frame2webs(family_groups, varnames = c("h_genus", "i_genus", "h_family"))
plotweb(family_web)
networklevel(family_web, index = "connectance")



```



### 1b. Do these analyses change if we look from the insect family perspective?
Make a more streamlined function to go through families 
```{r}
# Create df with families with fewer than 200 unique interactions
ins_data_new <- data %>% 
  group_by(i_genus) %>% 
  filter(i_genus == "Dendrocoris" | i_genus == "Myrmecoroides" | i_genus == "Xyonysius" | i_genus == "Anthelococcus" | i_genus == "Ceroplastes")
  # this pops out a df with 7619 observations

# Function that subsets each family and returns the shortened dataframe
ins_genus_data <- function(i) {
  igen <- subset(ins_data_new, ins_data_new$i_genus == i)
  return(igen)
}
insect_output<-list()

### before running this, all functions need to be rewritten

# new web function needs to be created
ins_web <- function(df) {
  as.data.frame(frame2webs(df, varnames = c("h_family", "i_genus", "i_genus"), type.out = 'array'))
}

i_genus_web <- ins_web(ins_data_new)

# multimetric should be edited to just pop out specialization metrics
species_multimet <- function(df) {
  c(
    # partner diversity
    part_div <- specieslevel(df, index = "partner diversity", level = "higher"),
    # proportional similarity
    prop_sim <- specieslevel(df, index = "proportional similarity", level = "higher"),
    # d
    d_spec <- specieslevel(df, index = "d", level = "higher")
  )
}

#### we might not need these old functions
for (i in unique_ifamilies) {
  ifam <- as.data.frame(ins_family_data(i))
  x <- multi_metric(web(ifam))
  y <- rbind(i, x)
  insect_output <- as.data.frame(rbind(output,y))
}
# FIX: printing y as a matrix rather than i attached to the values in x in one row


head(insect_output)

```


